Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime \thanks{ This work has been supported by the Italian PRIN Project COMMON-WEARS (2020HCWWLP) and the EU/MUR FSE REACT-EU PON R\&I 2014-2020. }

### 2023-05-21 paper-2023-acsos-pulverization.tex
intro

This kind of infrastructure is also valuable for  \acp{cas}~\cite{DBLP:journals/tpds/HongCHGZ19,DBLP:journals/comsur/WangZZMLW20,DBLP:journals/comsur/AfrinJRRWH21,IEEE-IoTJ-pulverization-simulation},
which generally feature components that,
depending on the conditions at hand,
could benefit from being deployed on different devices or from \emph{offloading} some of their  tasks.
%
However, exploiting this infrastructure,
 poses new challenges to \ac{cas} application designers.
%
Commonly, in fact, applications are designed with a specific infrastructure in mind,
whose assumptions unavoidably leak into the application logic~\cite{Spolsky2004},
unless captured and encapsulated away from it:
the higher the coupling between the application and the infrastructure,
the harder it is for the former to exploit the latter's full potential and adapt to changes.

The general solution is to devise 
 a reasonable \emph{partitioning}~\cite{DBLP:journals/jnca/LiuASGBQ15} of the software system
 and a corresponding (dynamic) deployment plan 
 defining the mapping between the software components and the target deployment domain~\cite{DBLP:journals/jss/ArcangeliBL15}.
%
In the context of \acp{cas},
 previous work proposes an application partitioning schema,
 called \emph{pulverisation}~\cite{FI2020-pulverization},
 the fosters the decoupling of business logic and deployment concerns.
%
In short, the core idea is to consider devices as \emph{logical} entities,
whose software can be designed as (or broken down to, if already existing) an ensemble of \emph{five components
(behaviour, state, communication, sensors, and actuators)}
 that can be deployed with flexibility on available infrastructure without (ideally) affecting application functionality. %whose actual host can be decided at deployment time.
%
This way, the designer can focus on the business logic at hand,
delaying an optimised deployment to later stages of the sofware design,
thus gaining resilience to changes in the infrastructure or in non-functional requirements.
%
One limitation of previous work on pulverisation regards dynamicity:
although the application can be deployed on arbitrary systems,
there is no support for the reconfiguration of components at runtime, as a reaction to changing conditions or requirements.

In this work,
we continue the research on pulverised systems by providing two key contributions:
\textbf{first}, we extend the theoretical framework of pulverisation by adding support for
\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
\textbf{second}, we provide a practical implementation in Kotlin called \emph{\ourframework{}} which,
to the best of our knowledge,
is the first \emph{practical framework supporting pulverisation}. % for a general-purpose language.
%
To evaluate \ourframework{}, we exercise it through two case studies:
a small-scale real-wold demonstration, and a city-scale simulated scenario.

>>>>

This kind of infrastructure is also valuable for  \acp{cas}~\cite{DBLP:journals/tpds/HongCHGZ19,DBLP:journals/comsur/WangZZMLW20,DBLP:journals/comsur/AfrinJRRWH21,IEEE-IoTJ-pulverization-simulation},
which generally feature components that,
depending on the conditions at hand,
could benefit from being deployed on different devices or from \emph{offloading} some of their  tasks.
%
However, exploiting this infrastructure
 poses new challenges to \ac{cas} application designers.
%
Commonly, in fact, applications are designed with a specific infrastructure in mind,
whose assumptions unavoidably leak into the application logic~\cite{Spolsky2004},
unless captured and encapsulated away from it:
the higher the coupling between the application and the infrastructure,
the harder it is for the former to exploit the latter's full potential and adapt to changes.

The general solution is to devise 
 a reasonable \emph{partitioning}~\cite{DBLP:journals/jnca/LiuASGBQ15} of the software system
 and a corresponding (dynamic) deployment plan 
 defining the mapping between the software components and the target deployment domain~\cite{DBLP:journals/jss/ArcangeliBL15}.
%
In the context of \acp{cas},
 previous work proposes an application partitioning schema,
 called \emph{pulverisation}~\cite{FI2020-pulverization},
 that fosters the decoupling of business logic and deployment concerns.
%
In short, the core idea is to consider devices as \emph{logical} entities,
whose software can be designed as (or broken down to, if already existing) an ensemble of \emph{five components
(behaviour, state, communication, sensors, and actuators)}
 that can be deployed with flexibility on available infrastructure without (ideally) affecting application functionality. %whose actual host can be decided at deployment time.
%
This way, the designer can focus on the business logic at hand,
delaying an optimised deployment to later stages of the sofware design,
thus gaining resilience to changes in the infrastructure or in non-functional requirements.
%
One limitation of previous work on pulverisation regards dynamicity:
although the application can be deployed on arbitrary systems,
there is no support for the reconfiguration of components at runtime, as a reaction to changing conditions or requirements.

In this work,
we further develop the research on pulverised systems by providing two key contributions:
%
\begin{enumerate}
 \item we extend the theoretical framework of pulverisation by adding support for
\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
 \item we accordigly provide a practical implementation of pulverisation (with reconfiguration rules) called \emph{\ourframework{}}, developed in the Kotlin programming language.
\end{enumerate}
%
To evaluate \ourframework{}, we exercise it through two case studies:
a small-scale real-wold demonstration, and a city-scale simulated scenario.
